Measuring Perceptions and Preferences for Meritocracy

### 2020-07-30 docs/paper.Rmd
likert, corplot, tablefit, sempath, bib, preanalisis

La desigualdad económica se ha vuelto un tema que genera creciente preocupación y malestar alrededor del mundo. Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@PikettyCapitalTwentyFirstCentury2014b]. En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_Fairness_2005; @dimick_Models_2018] la legitimación de la desigualdad económica [@schroder_Income_2017]  y el funcionamiento de la meritocracia [@Duru-BellatWhoMeritocracyIndividual2012b; @MijsParadoxInequalityIncome2019b; @ReynoldsPerceptionsMeritocracyLand2014b].

En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962]. Esta concepción tradicional de mérito pone en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994]. Una serie de estudios han realizado críticas a la realización de este estándar moral de distribución, planteando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@WitteveenReconsideringmeritocraticpower2020a; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019]. Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@PremingerMeritocracyserviceethnocracy2020; @TrumpWhenwhyeconomic2020; @Meredithsocietydivisibleblessed2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].

Debido al rol que cumplen las creencias meritocráticas dentro del pensamiento neoliberal [@Bay-ChengTrackingHomoOeconomicus2015], han surgido múltiples investigaciones que evalúan la relación entre creencias meritocráticas y diversos ámbitos sociales de la actualidad. Por ejemplo, se han desarrollado estudios que vinculan la meritocracia al reforzamiento de estereotipos socioeconómicos, de género y de etnias [@MadeiraPrimesConsequencesSystematic2019; @GirerdNeoliberalismIdeologicalBarrier2020; @PremingerMeritocracyserviceethnocracy2020], así como también líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo [@GenerettStoriesWeTell2020; @ReayPerilsPenaltiesMeritocracy2020; @OwensENGINESSOCIALMOBILITY2020] y en el contexto organizacional de las empresas [@PerezAdvancingcareersmerit2020; @AielloNewevidencefirmuniversity2019].

@@

In Mijs' proposal we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization). The problem of whether talent is or not meritocratic is something that actually should (and can) be tested, it cannot be simply ruled out particularly in empirical research. This conceptual and measurement issue it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@MaitnerEmotionalreactionsunequal2015;@MajorPerceiveddiscriminationworldview2007; @BubakPerceptionsmeritocracynote2019; @solt_economic_2016].

@@

The "resons to get ahead" battery refers to "how important you think it is", considered by Mijs as a belief in meritocracy for the item regarding effort. Nevertheless, another version of this same battery in the survey asks about "how important you think it _should_ be".  Which one of both is a "belief"? According to @janmaat_subjective_2013-1 : "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Therefore, the referred paper would be actually using the term beliefs referring to perceptions. The same occurs in @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use different terms as attitudes (@kunovich_systems_2007). The first attempt to shed light on this confusion was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “... people are rewarded for their skills?”.

@@

Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brown_confirmatory_2006, @bollen_structural_1989], using single indicators and/or simple average indexes for measuring meritocracy. Such strategy assumes that the construct is measured perfectly by the indicators chosen, going as far as proposing that "... In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654]. Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily they end up in a sum index despite proving multidimensionality.

@@

The study was pre-registered ... (link) ***

>>>>

La desigualdad económica se ha vuelto un tema que genera creciente preocupación y malestar alrededor del mundo. Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@piketty_Capital_2014]. En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_Fairness_2005; @dimick_Models_2018] la legitimación de la desigualdad económica [@schroder_Income_2017]  y el funcionamiento de la meritocracia [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014].

En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962]. Esta concepción tradicional de mérito pone en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994]. Una serie de estudios han planteado críticas a la realización de este estándar moral de distribución, argumentando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019]. Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].

Debido al rol que cumplen las creencias meritocráticas dentro del pensamiento neoliberal [@bay-chengTrackingHomoOeconomicus2015], han surgido múltiples investigaciones que evalúan la relación entre creencias meritocráticas y diversos ámbitos sociales de la actualidad. Por ejemplo, se han desarrollado estudios que vinculan la meritocracia al reforzamiento de estereotipos socioeconómicos, de género y de etnias [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], así como también líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] y en el contexto organizacional de las empresas [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].

@@

In Mijs' proposal we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization). The problem of whether talent is or not meritocratic is something that actually should (and can) be tested, it cannot be simply ruled out particularly in empirical research. This conceptual and measurement issue it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].

@@

The "resons to get ahead" battery refers to "how important you think it is", considered by Mijs as a belief in meritocracy for the item regarding effort. Nevertheless, another version of this same battery in the survey asks about "how important you think it _should_ be".  Which one of both is a "belief"? According to @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Therefore, the referred paper would be actually using the term beliefs referring to perceptions. The same occurs in @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use different terms as attitudes (@kunovich_systems_2007). The first attempt to shed light on this confusion was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “... people are rewarded for their skills?”.

@@

Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989], using single indicators and/or simple average indexes for measuring meritocracy. Such strategy assumes that the construct is measured perfectly by the indicators chosen, going as far as proposing that "... In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654]. Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily they end up in a sum index despite proving multidimensionality.

@@

A pre-registration was made in the OSF platform, [Perceptions and Preferences for Meritocracy Scale
Public registration](https://osf.io/z45y2) In it, it was indicated, before carrying out the analyzes, the four-dimensional theoretical model, the level of measurement of the variables, the statistical tests to be used with their respective evaluation parameters, along with other important aspects of the research design.
